Weather Data Integration and Assimilation System\\

### 2020-02-21 introduction/sec_1-motivation.tex
Finished fews citations

Weather prediction is essential to reduce impact due to natural disasters and to gain effectively manage natural resources like water. It will give more understanding of weather prediction if we look into a weather prediction flow that can use to predict the surface water level. The weather predictions/forecasts are performed using \acrfull{nwm} that use a set of configuration parameters and the current status of the environment such as temperature, rainfall, and wind direction as the input. For example, the \acrfull{wrf} \cite{MesoscaleMicroscaleMeteorologyLaboratoryWeatherModel} model is used to predict the rainfall, temperature, and wind direction. Typically multiple \acrshort{nwm} are executed on the same data with different configuration parameters. Then using the real-time observations, the correct model is selected. After removing the bias, the extract information is then fed into another set of models/simulators. 
As an example, FLO2D is a hydrologic model that can use to predict the surface water level. A given FLO2D model can simulate with the initial surface water levels and the sequence of forecasted rainfall for the corresponding modeled area from the \acrshort{nwm} model to forecast future water levels.
\dbc{Highighted part doesn't make much sense to me. Need to be reworded.}
\gkc{Highlighted parts modified and kept also.}
\dbc{cite both NWM and WRF}
\gkc{FIXED BOTH}

\acrfull{curw} \cite{CUrWSL2017SL} is tasked with the mission to reduce water-related natural disasters in Sri Lanka. \acrshort{curw} \cite{CUrWSLObservedSL} is implementing its system with the idea of long-term maintainability rather depend on existing weather data integration systems. It collected real-time data from weather stations and other government entities and generates daily forecasts. Then the generated results are shared with relevant public departments responsible for disaster planning, mitigation, and management. One of the future goals of \acrshort{curw} is to open up data to the public enabling both societal and commercial use cases.

@@

Processed weather data needs to be accessed by many third-party applications and research to make use of them. For example, logistics companies could use those data with their models to plan and schedule their deliveries. Agricultural insurance companies can warn the farmers, as well as calculate premiums based on anticipated weather patterns. Also, farmers rely on weather forecasts to decide what work to do on any particular day. For example, drying hay is only feasible in dry weather. Further, the aviation industry is especially sensitive to the weather, accurate weather forecasting is essential since before takeoff on the conditions to expect en route and at their destination. 
Thus it is required a platform to provide such 3rd party access, and use the data or process those based on their applications.

While Data Integration and Analysis System (DIAS) \cite{Kawasaki2018DataReduction} and Meteorological Assimilation Data Ingest System (MADIS) \cite{Macdermaid2005ARCHITECTUREP2.39} are some of well-known Weather Data Integration and Assimilation (WDIA) systems, they are proprietary. Further \acrfull{fews} \cite{Werner2013TheSystem} is free to use, it is not open source.
Even though there are many data integration systems, most of them are proprietary or closed source. For an island like Sri Lanka that having different kinds of weather seasons over the year than that software originated need to be highly customized. Throughout the operations and maintenance, it is required to get support with paying lots of charges, and difficult to debug into a closed source system without maintenance support from the vendors. As an example, \acrshort{fews} does not support FLO2D model integration, and \acrshort{curw} is using FLO2D for water level forecasting in the data flow. In such a case, it required to get support from the vendor to integrate and configure the system.
And most of the software existing out there is not up to date with the latest technology concepts such as Cloud Computing. The modern architecture pattern such as microservice architecture was able to provide high scalability and availability, rather using monolithic architecture or client service architecture. Most existing systems required to shut down to update the system configuration such as \acrshort{fews}, but a system that is accessed by any third parties should be operating 99.9\% without any downtime and should allow adding new features on the fly with rollover capabilities. The proposed system should be independent of the platform. But most software out there is not, as an example \acrshort{fews} can run on Windows and Linux. And should be independent underline infrastructure providers which gives much flexibility for the users to decide the best cost-effective way to host it. Given the above facts, there is a necessity for such a system to be implemented.

>>>>

Weather prediction is essential to reduce the impact caused by natural disasters and to gain effectively manage natural resources like water. It will give more understanding of weather prediction if we look into a weather prediction flow that can use to predict the surface water level. The weather predictions/forecasts are performed using \acrfull{nwm} that use a set of configuration parameters and the current status of the environment such as temperature, rainfall, and wind direction as the inputs. For example, the \acrfull{wrf} \cite{MesoscaleMicroscaleMeteorologyLaboratoryWeatherModel} model is used to predict the rainfall, temperature, and wind direction. Typically multiple \acrshort{nwm} are executed on the same data with different configuration parameters. Then using the real-time observations, the correct model is selected. After removing the bias, the extract information is then fed into another set of models/simulators. 
\db{As an example, FLO2D is a hydrologic model that can use to predict the surface water level. A given FLO2D model can simulate with the initial surface water levels and the sequence of forecasted rainfall for the corresponding modeled area from the \acrshort{nwm} model to forecast future water levels.}
\dbc{Highighted part doesn't make much sense to me. Need to be reworded.}
\gkc{Highlighted parts modified and kept also.}
\dbc{cite both NWM and WRF}
\gkc{FIXED BOTH}

\acrfull{curw} \cite{CUrWSL2017SL} is tasked with the mission to reduce water-related natural disasters in Sri Lanka. \acrshort{curw} \cite{CUrWSLObservedSL} is implementing own system with the idea of long-term maintainability rather depend on existing weather data integration systems. It collected real-time data from weather stations and other government entities and generates daily forecasts. Then the generated results are shared with relevant public departments responsible for disaster planning, mitigation, and management. One of the future goals of \acrshort{curw} is to open up data to the public enabling both societal and commercial use cases.

@@

Processed weather data needs to be accessed by many third-party applications and research to make use of them. For example, logistics companies could use those data with their models to plan and schedule their deliveries. Agricultural insurance companies can warn the farmers, as well as calculate premiums based on anticipated weather patterns. Besides, farmers rely on weather forecasts to decide what work to do on a given day. For example, drying hay is only possible in dry weather. Also, the aeronautical industry is particularly sensitive to weather conditions, precise weather forecasts are essential since before takeoff a flight need to check the conditions on expected route and at destination.
Thus it is required a platform to provide such 3rd party access, and use the data or process those based on their applications.

While Data Integration and Analysis System (DIAS) \cite{Kawasaki2018DataReduction} and Meteorological Assimilation Data Ingest System (MADIS) \cite{Macdermaid2005ARCHITECTUREP2.39} are some of well-known Weather Data Integration and Assimilation (WDIA) systems, they are proprietary. Further \acrfull{fews} \cite{Werner2013TheSystem} is free to use, it is not open source.
Even though there are many data integration systems, most of them are proprietary or closed source. For an island like Sri Lanka that having different kinds of weather seasons over the year than that software originated need to be highly customized. Throughout the operations and maintenance, it is required to get support with paying lots of charges, and difficult to debug into a closed source system without maintenance support from the vendors. As an example, \acrshort{fews} does not support FLO2D model integration, and \acrshort{curw} is using FLO2D for water level forecasting in the data flow. In such a case, it required to get support from the vendor to integrate and configure the system.

Most of the software existing out there is not up to date with the latest technology concepts such as Cloud Computing. The modern architecture pattern such as microservice architecture was able to provide high scalability and availability, rather using monolithic architecture or client service architecture. Most existing systems required to shut down to update the system configuration such as \acrshort{fews}, but a system that is accessed by any third parties should be operating 99.9\% without any downtime and should allow adding new features on the fly with rollover capabilities. The proposed system should be independent of the platform. But most software out there is platform dependent, as an example \acrshort{fews} can run on Windows and Linux. And should be independent underline infrastructure providers which gives much flexibility for the users to decide the best cost-effective way to host it. Given the above facts, there is a necessity for such a system to be implemented.

### 2020-02-21 introduction/sec_2-problem_and_objectives.tex
Finished fews citations

\section{Problem Statement}
Handling Spatio-temporal weather data is challenging \dbc{as} it involves large data volumes required to have a higher number of resources such as computation, storage, and network bandwidth. Moreover, the data integration system needs to be capable of interacting with different data formats including Spatio-temporal bulk stream data. Furthermore, it should support geographical and time-based queries, as well as provide third-party access to data with easy integration. Scalability and high availability are also essential to support different use cases and mission-critical nature of applications.
And users should be able to easily deploy and manage the system with modern technologies and should be able to get the advantage of technology advancement. In this context, the research problem can be stated as follows:

\emph{Design and implement an extendable weather data integration, assimilation, and dissemination system which is capable of handling bulk data efficiently with providing scalability and high throughput with using open source tools and modern cloud computing technologies.}
\dbc{Give a proper problem statement.}
\gkc{UPDATED. Looked into the OneNote samples. I  should follow one of them from the beginning. :( }

\section{Objectives}
We address the above problem statement by achieving the following research objectives:
\begin{itemize}
    \item To develop an open platform to integrate weather data from different sources in different formats
    \item To design and develop a schema to store multidimensional weather timeseries data while optimizing access time, availability, and storage
    \item To optimize schema to support data queries based on geography and time
    \item To develop an API for third-party data sharing
%    \item To use open source tools and provide the source publicly with appropriate liecense.
\end{itemize}

>>>>

\section{Problem Statement}
Handling Spatio-temporal weather data is challenging \db{as} it involves large data volumes required to have a higher number of resources such as computation, storage, and network bandwidth. Moreover, the data integration system needs to be capable of interacting with different data formats including Spatio-temporal bulk stream data. Furthermore, it should support geographical and time-based queries, as well as provide third-party access to data with easy integration. Scalability and high availability are also essential to support different use cases and mission-critical nature of applications.
And users should be able to easily deploy and manage the system with modern technologies and should be able to get the advantage of technology advancement. In this context, the research problem can be stated as follows:

\emph{Design and implement an extendable weather data integration, assimilation, and dissemination system which is capable of handling bulk data efficiently with providing scalability and high throughput with using open source tools which support modern cloud computing advancements.}
\dbc{Give a proper problem statement.}
\gkc{UPDATED. Looked into the OneNote samples. I  should follow one of them from the beginning. :( }

\section{Objectives}
This research addresses the above mentioned problem statement by achieving the following research objectives:
\begin{itemize}
    \item To develop an open platform to integrate weather data from different sources in different formats
    \item To design and develop a schema to store multidimensional weather timeseries data while optimizing access time, availability, and storage
    \item To optimize schema to support data queries based on geography and time
    \item To develop an API for third-party data sharing
%    \item To use open source tools and provide the source publicly with appropriate liecense.
\end{itemize}

### 2020-02-21 introduction/sec_3-outline.tex
Finished fews citations

\section{Outline}
The rest of the thesis is organized as follows. The literature review is presented in \cref{ch:literature}, and discuss the existing Weather Data Integration Systems such as \acrshort{lead}, \acrshort{fews}, \acrshort{madis}, and \acrshort{dias}. Solution approach with the proposed system architectures and adaptation with problem occurred are presented in Chapter \ref{ch:method}. It provides details about how the \acrshort{wdias} is designed by following a microservice architecture and the data preprocessing capabilities of the system. Performance analysis is presented in Chapter \ref{ch:results}. Here we discuss the experimental setup of the \acrshort{wdias} and analyze the observations concerning performance metrics. Finally, in Chapter \ref{ch:summary}, we conclude and summarize the work and further discussed the research limitations and future work.

>>>>

\section{Outline}
The organization of the rest of the thesis as follows. In next \cref{ch:literature} the literature review is present and discuss the existing Weather Data Integration Systems such as \acrshort{lead}, \acrshort{fews}, \acrshort{madis}, and \acrshort{dias}. Solution approach with the proposed system architectures and adaptation with problem occurred are presented in Chapter \ref{ch:method}. It provides details about how the \acrshort{wdias} is designed by following a microservice architecture and the data preprocessing capabilities of the system. Performance analysis is presented in Chapter \ref{ch:results}. Here we discuss the experimental setup of the \acrshort{wdias} and analyze the observations concerning performance metrics. Finally, in Chapter \ref{ch:summary}, we conclude and summarize the work and further discussed the research limitations and future work.

### 2020-02-21 lit/sec_1-fews.tex
Finished fews citations

\acrshort{fews} \cite{Werner2013TheSystem} is developed by Deltares in the Netherland with targeted for use by operational forecasting agencies. The codebase of \acrshort{fews} is currently not fully open source.
Most, if the models are using a model-centric approach such as inputs, need to be in the format of specific to the model. Also, the outputs produced by the model are specific to the model and hard to utilize by other systems (e.g., FLO2D). As \acrshort{fews} uses a modular approach it is easier to integrate new models. 
Thus, \acrshort{fews} can be considered as an integration framework or a middleware for the models.

Forecasting processes are constructed as a combination of modelling steps and data transformation algorithms. These are then combined to the provide required forecast capabilities. Flexibility is achieved by integrating new models and algorithms into the codebase \cite{Werner2013TheSystem}. \acrshort{fews} system contains no inherent hydrological modeling capabilities within its codebase. Rather it is a framework that can use to integrate into the system and create workflows for forecasting.

Are proposed by Haggett \cite{Haggett1998AnWales} key elements of a forecasting system include detection, forecasting, dissemination and warning, and response. Within these four steps, \acrshort{fews} focuses on the forecasting step. The primary objective of this step is to provide additional lead time through predictions of short term future hydro-meteorological conditions \cite{Werner2005FloodCatchments}. It is a valid argument that providing accurate predictions with greater lead time can reduce the level of destruction. To do the forecasting, the system should be capable of integrating real-time data from hydrological and meteorological observation networks, and the dissemination of prediction results through appropriate products to the warning process.

In \db{Figure} \cref{fi:fews_schematic} show a schematic view of the connection between the forecasting system to real-time data acquisition systems and dissemination systems. The widely used concept is using meteorological forecast data to get precipitation and, then using a hydrological and hydraulic model chain to predict the level affect on the ground. The hydrological and hydraulic model should be design based on the Geographical data. The ground should be analyzed geographically and divide into the catchments. Based on the affected catchment, it can be further divided into sub-catchments to reduce the complexity of the simulation task. Ideally, the forecasting system should be flexible to allow change to models and data, while keeping the way forecasters work with it as constant as possible.
\dbc{F is capital and T is capital when referring to Figures and Tables. Fix all other places.}
\gkc{FIXED}

\begin{figure}[htp]
    \centering
    \includegraphics[width=0.7\textwidth]{lit/fews/Schematic-structure-of-a-fl-ood-forecasting-system-showing-the-position-of-Delft-FEWS_W640.png}
    \caption[Schematic structure of a flood forecasting system including \acrshort{fews} and links to other primary systems within the operational environment]{Schematic structure of a flood forecasting system including \acrshort{fews} and links to other primary systems within the operational environment \cite{Werner2013TheSystem}.}
    \label{fi:fews_schematic}
\end{figure}

The foundation of \acrshort{fews} is data-centric, with a common data-model through which all components interact. All timeseries data (both scalar and gridded) are stored in this common data model in a database. Modeling capabilities are then linked to the system through one of the interfaces provided to the data-model \cite{Werner2013TheSystem}. As mention in the paper, having a common data-model make easy to store data efficiently. Operations like reporting and sharing the data also make it easy by integrating models. But the problem with this approach is, handling the multiple data formats. \acrshort{fews} overcome this issue by introducing adapters for most of the commonly used data formats. This is a good feature that enables users to easily import data into the system. But at the same time, it's adding additional complexity to the system.

Timeseries data is additionally either scalar, vector, or gridded data, though all different types are uniformly stored as binary objects in a time series table. None of the functional components (including the linked models) have direct access to the times series table, which is accessible only through the data access module \cite{Werner2013TheSystem}. As explained in the system interpretation, every time all the data access needs to go through the data access module which causing adding more cost in data access. Since it is storing the different data types as binary objects, there is a penalty in converting data into binary objects and vice versa. Storing all the data in a timeseries table causing all the requests to come into a single data model in a database. Even it gives the advantage of accessing all the data in one place. Since screaming big amount of data is causing a heavy load on the system, causing to slow down in the performance of trying to use the system on a large set of data.

Given above drawback on storing the data in the system, within the \acrshort{fews} data model, time series are uniquely identified by their location and data type, as well as an id related to the source of the data (e.g., the external source or hydrological model of which the time series are a result) \cite{Werner2013TheSystem}. This allows indexing the database on the above key fields and stores the data separately, to take advantage of putting multiple data resources into the system. As an example, instead of storing the data at a single timeseries table, it is possible to separate and index the database based on the source such as an external source or hydrological model of which the timeseries is a result. Or separate by data type and store in multiple storages give the capability for the system to scale with a factor of identical types that can be identified. Example of separation by data types of scalar, vector and gridded will increase the scaling factor by 3x.

Data processing and manipulation is a required process in weather forecasting. Most of the data that is imported from external sources is not at the appropriate temporal and spatial scale to be applied as an input to a forecasting model, or to be used directly in product generation. As a consequence, generic data processing steps form the predominant effort in most applications of models in the forecast environment. Some examples include data validation, serial and spatial interpolation, aggregation and disaggregation, and merging data \cite{Werner2013TheSystem}. This a vital feature in a forecasting system, and affect the quality and accuracy of the predicted data outputs. Because of the common data model concept in \acrshort{fews}, data processing via these functions is much effective. The system is self-provided some of the generic functions for data processing, but for more complex algorithms can be developed as a new Java class coded to communicate with the \acrfull{api}. It is obvious that for additional feature integration, users should implement the new functional extensions via Java programming language. Users do not have the flexibility of developement with some other languages they are familiar with or use the support from another language. As an example, Python language is easy to use for beginners and many data scientists are using this programming language for development and it has many data available libraries for data processing. But in \acrshort{fews}, users cannot take advantage of such existing tools.
\dbc{Define Abb. for API in Chapter 1.}
\gkc{FIXED}

\begin{figure}[htp]
    \centering
    \includegraphics[width=1.0\textwidth]{lit/fews/Architecture-of-Delft-FEWS-showing-the-data-base-the-data-access-layers-and-examples-of_W640.png}
    \caption[Architecture of \acrshort{fews}]{Architecture of \acrshort{fews} \cite{Werner2013TheSystem}.}
    \label{fi:fews_data_layer}
\end{figure}
Most of the library of functions provided work equally on scalar and gridded time series. As with all other modules, this communicates with the database solely through the data access layer as shown in \cref{fi:fews_data_layer} \cite{Werner2013TheSystem}. As it is shown in the figure, it is clearly showing that the system depends on a single database and via the data access layer, all the requests are coming to the database through a connection. It is possible to setup and connect to an enterprise-level database with clustering and shading as a paid solution, to serve many requests as possible. But the design itself inherently suffers from the database bottleneck which effected by running a large set of timeseries or models.
\dbc{When referring to a section you need to use the word "Section" as highlighted above. Fix all other places.}
\gkc{FIXED. I did not remove highlited places as you need to verify.}

\begin{figure}[htp]
    \centering
    \includegraphics[width=0.8\textwidth]{lit/fews/Linking-Delft-FEWS-with-external-models-The-fi-gure-shows-the-fl-ow-of-data-through-XML_W640.png}
    \caption[Linking \acrshort{fews} with external models]{Linking \acrshort{fews} with external models \cite{Werner2013TheSystem}.}
    \label{fi:fews_general_adapter}
\end{figure}
One of the simple and most effective features in the \acrshort{fews} is an open approach to the integration of models and data. This approaches the concept of the open modeling framework proposed by Open model integration \cite{Kokkonen2003InterfacingXML}. It simply gives the flexibility to allow operators to integrate more models as well as variations of the same model and come up with new integration flows as much as possible.
\acrshort{fews} generates the input data as a set of XML files to a defined location; an adapter developed specifically for the model in question transforms this to the required native format in a pre-processing step; \acrshort{fews} executes the model, and the adapter to that model then converts the native formatted results into XML formatted files in a post processing step. \acrshort{fews} subsequently imports the results into the database from the XML files (see \cref{fi:fews_general_adapter}) \cite{Werner2013TheSystem}. The process is simple and consistent throughout all the model integration. The model execution is done by the \acrshort{fews} or the model adapter which is causing tide coupling into the execution process in the system. Users do not have the flexibility to run the models with different configurations such as running with parallel execution. That may be overcome by triggering an external process at the execution time, but it introduces more difficulty for handling after model executed successfully, preceding to the next step in the process. This part of the system feature is not focused on the \acrshort{wdias} and user has to come up with own flow or use existing scientific flow management system. But it is a concept that needs to be discussed and understand properly to design a data integration and assimilation system. But the \acrshort{wdias} is capable of integrating workflow mechanism with its extendable architecture.

Exchanging data with the model is primarily through XML files. In some cases, these XML files may become very large, which may lead to I/O bottlenecks and subsequent performance issues \cite{Werner2013TheSystem}. This issue is thoroughly discussed in previous paragraphs and the authors of the \acrshort{fews} \cite{Werner2013TheSystem} seem to be noticed and accept it here. For overcoming this issue they introduced the file-based exchange of data. This includes the use of binary-XML files, streaming files through memory, and the use of \acrshort{netCDF} files. But far as for the users, it seems to be adding more complexity and context to get higher performance via the system.

If a particular adapter for a model is not present in the list of \acrshort{fews} adapter, the user has the flexibility to implement an adapter for it. The effort of developing an adapter for a model code not previously integrated with \acrshort{fews} will vary depending on the complexity of the model I/O formats \cite{Werner2013TheSystem}. If the users want to use an existing adapter and need to configure something not supported via the existing adapter, it is hard to get done. Also, there are some cases it is hard to develop an adapter or running alongside the \acrshort{fews}. As an example, the FLO2D model is using for hydrologic modeling and the model only supports only on Windows based operating systems. If you setup the system on Linux based operating system and using Linux based models, it is hard for users to come up with a solution for integrating the FLO2D models.
The final step and one of important feature is exporting the data from the system in useful manner.
The forecasting process is often a sequence of steps, starting with the import of data, a several data processing, and modeling steps, and culminating in the generation of products to be disseminated to the warning process. This is supported via the workflow process in \acrshort{fews}. As mentioned above, the scope of the \acrshort{wdias} is not focused on adding such a feature in the system, rather user has to use a scientific workflow management system or come up with own version of workflow.

When running sequential steps within a workflow in \acrshort{fews}, each step is run for the full-time window required (e.g., the lead time of the forecast) before moving on to the next step. This precludes tightly coupling of models, even explicitly \cite{Werner2013TheSystem}. As mentioned above, users can run multiple models parallel or run those sequentially as needed.

>>>>

\acrshort{fews} \cite{Werner2013TheSystem} is developed by Deltares in the Netherland with target audiences/customers of operational forecasting agencies. The \acrshort{fews} codebase is not fully open source at the moment.
Most of models are using a model-centric approach such as inputs, need to be in the format of specific to the model. Also, the outputs produced by the model are specific to the model and hard to utilize by other systems (e.g., FLO2D create human readable text output files). As \acrshort{fews} uses a modular approach it is easier to integrate new models. 
Thus, \acrshort{fews} can be considered as an integration framework or a middleware for the models.

As it describe with a simple example in \cref{ch:intro}, forecasting processes consist by creating modeling steps with combining data transformation algorithms. Combining such flows can use to provide the forecasting capabilities. The system is flexible enough to integrate new models and algorithms into the core codebase \cite{Werner2013TheSystem}. \acrshort{fews} does not have any inbuilt hydrological modeling capabilities within its codebase. Rather it is a framework that can use to integrate into the system and create workflows for forecasting.

Are proposed by Haggett \cite{Haggett1998AnWales} key elements of a forecasting system are defined as detection, forecasting, then dissemination and warning, and response to the events. The \acrshort{fews} focuses on the forecasting step out from above steps. The main objective of this step is to provide additional lead time by forecasting future hydro-meteorological conditions \cite{Werner2005FloodCatchments}. It is a valid argument that providing accurate predictions with greater lead time can reduce the level of destruction. To forecasting, the system should be capable of integrating real-time data from meteorological and hydrological observation network systems, and the disseminate the forecasted results through relevant products to the warning process.

In \db{Figure} \cref{fi:fews_schematic} show the connection between the forecasting system to real-time data integration systems and dissemination systems. The widely used concept is using meteorological forecast data to get precipitation and, then using a hydrological and hydraulic model chain to predict the level affect on the ground. The hydrological and hydraulic model should be design based on the Geographical data. The ground should be analyzed geographically and divide into the catchments. Based on the affected catchment, it can be further divided into sub-catchments to reduce the complexity of the simulation task. Ideally, the forecasting system should be flexible enough to allow modification of models and data, while keeping the most constant way possible of working with forecasters.
\dbc{F is capital and T is capital when referring to Figures and Tables. Fix all other places.}
\gkc{FIXED}

\begin{figure}[htp]
    \centering
    \includegraphics[width=0.7\textwidth]{lit/fews/Schematic-structure-of-a-fl-ood-forecasting-system-showing-the-position-of-Delft-FEWS_W640.png}
    \caption[Schematic structure of a flood forecasting system including \acrshort{fews} and communication between other operational systems]{Schematic structure of a flood forecasting system including \acrshort{fews} and communication between other operational systems \cite{Werner2013TheSystem}.}
    \label{fi:fews_schematic}
\end{figure}

The \acrshort{fews} follows a data-centric approach, with providing a common data-model to interact from all other components. All the timeseries data types are stored in a database which follows a common data model. New models are integrate to the system via using a one of the interfaces provided to interact with the common data-model \cite{Werner2013TheSystem}. As mention in the paper, having a common data-model make easy to store data efficiently. Operations like reporting and sharing the data also make it easy by integrating models. But the problem with this approach is, handling the multiple data formats. \acrshort{fews} overcome this issue by introducing adapters for most of the commonly used data formats. This is a good feature that enables users to easily import data into the system. But at the same time, it's adding additional complexity to the system.

The timeseries data is one of scalar, vector, or grid data, and all of different types are uniformly stored as binary objects in a time series table inside the \acrshort{fews}. Functional compenents or any integrated models do not have direct access to the timeseries table which mentioned above, and those should use the data access module to access the data \cite{Werner2013TheSystem}. As explained in the system interpretation, every time all the data access needs to go through the data access module which causing adding more cost in data access. Since it is storing the different data types as binary objects, there is a penalty in converting data into binary objects and vice versa. Storing all the data in a timeseries table causing all the requests to come into a single data model in a database. Even it gives the advantage of accessing all the data in one place. Since screaming big amount of data is causing a heavy load on the system, causing to slow down in the performance of trying to use the system on a large set of data.

Given above drawback on storing the data in the system, the \acrshort{fews} data model store the timeseries by uniquely identifying based on the location, data type, and source of the data \cite{Werner2013TheSystem}. This allows indexing the database on the above key fields and stores the data separately, to take advantage of putting multiple data resources into the system. As an example, instead of storing the data at a single timeseries table, it is possible to separate and index the database based on the source such as an external source of data or hydrological model of which the timeseries is a result. Or separate by data type and store in multiple storages give the capability for the system to scale with a factor of identical types that can be identified. Example of separation by data types of scalar, vector and gridded will increase the scaling factor by 3x.

Data processing and manipulation is a required process in weather forecasting. 
The data integrated from the external sources are not in the appropriate temporal and spatial format that can directly feed as an input to a forecasting model or use in other application. Therefore, generic data processing steps are a required effort in most model applications in the forecasting environment. Few examples are serial and spatial interpolation, data validation, aggregation and disaggregation, and data fusion \cite{Werner2013TheSystem}. This a vital feature in a forecasting system, and affect the quality and accuracy of the predicted data outputs. Because of the common data model concept in \acrshort{fews}, data processing via these functions is much effective. The system is self-provided some of the generic functions for data processing, but required complex algorithms can be implement as a new Java class via communicate with the \acrfull{api}. It is obvious that for additional feature integration, users should implement the new functional extensions via Java programming language. Users do not have the flexibility of development with some other languages they are familiar with or use the support from another language. As an example, Python language is easy to use for beginners and many data scientists are using this programming language for development and it has many data available libraries for data processing. But in \acrshort{fews}, users cannot take advantage of such existing tools.
\dbc{Define Abb. for API in Chapter 1.}
\gkc{FIXED}

\begin{figure}[htp]
    \centering
    \includegraphics[width=1.0\textwidth]{lit/fews/Architecture-of-Delft-FEWS-showing-the-data-base-the-data-access-layers-and-examples-of_W640.png}
    \caption[Architecture of \acrshort{fews}]{Architecture of \acrshort{fews} \cite{Werner2013TheSystem}.}
    \label{fi:fews_data_layer}
\end{figure}
Most of the function library has provided equal work on scalar and grid time series. But the communicates with the database only via the data access layer as shown in \cref{fi:fews_data_layer} \cite{Werner2013TheSystem}. As it is shown in the figure, it is clearly showing that the system depends on a single database and via the data access layer, all the requests are coming to the database through a connection. It is possible to setup and connect to an enterprise-level database with clustering and shading as a paid solution, to serve many requests as possible. But the design itself inherently suffers from the database bottleneck which effected by running a large set of timeseries or models.
\dbc{When referring to a section you need to use the word "Section" as highlighted above. Fix all other places.}
\gkc{FIXED. I did not remove highlited places as you need to verify.}

\begin{figure}[htp]
    \centering
    \includegraphics[width=0.8\textwidth]{lit/fews/Linking-Delft-FEWS-with-external-models-The-fi-gure-shows-the-fl-ow-of-data-through-XML_W640.png}
    \caption[\acrshort{fews} integration with external models]{\acrshort{fews} integration with external models \cite{Werner2013TheSystem}.}
    \label{fi:fews_general_adapter}
\end{figure}
One of the simple and most effective features in the \acrshort{fews} is an open approach to the integration of models and data. The concept of the open modeling framework is used by the \acrshort{fews} which is proposed by Open model integration \cite{Kokkonen2003InterfacingXML}. It simply gives the flexibility to allow operators to integrate more models as well as variations of the same model and come up with new integration flows as much as possible.
The users need to mentioned the input timeseries via XML configurations, and \acrshort{fews} generates the input data as XML files. Then using an adapter developed for the model transforms the data into required format for the model as a preprocessing step. Next \acrshort{fews} executes the model and model generates the output after successful completion. In the post processing step, the model result data transform to the XML files which can import into the data model as shown in \cref{fi:fews_general_adapter} \cite{Werner2013TheSystem}. The process is simple and consistent throughout all the model integration. The model execute by the \acrshort{fews} or the model adapter which is causing tide coupling into the execution process in the system. Users do not have the flexibility to run the models with different configurations such as running with parallel execution. That may be overcome by triggering an external process at the execution time, but it introduces more difficulty for handling after model executed successfully, preceding to the next step in the process. This part of the system feature is not focused on the \acrshort{wdias} and user has to come up with own flow or use existing scientific flow management system. But it is a concept that needs to be discussed and understand properly to design a data integration and assimilation system. But the \acrshort{wdias} is capable of integrating workflow mechanism with its extendable architecture.

Data exchange with the model is mainly done via XML files. It is possible to happen, these XML files become very large, which can lead to I/O bottlenecks and causing performance issues \cite{Werner2013TheSystem}. This issue is thoroughly discussed in previous paragraphs and the authors of the \acrshort{fews} \cite{Werner2013TheSystem} seem to be noticed and accept it here. For overcoming this issue they introduced the file-based exchange of data. This includes the use of binary XML files, the streaming of files via memory and the use of \acrshort{netCDF} files. But far as for the users, it seems to be adding more complexity and context to get higher performance via the system.

The \acrshort{fews} provides a list of adapter available. If an adapter is not available for a given model, the user has to implement the adapter for the model and \acrshort{fews} has the flexibility creating a new adapter. But the effort that required to develop an adapter for a model will vary depending on the complexity of the model required data formats \cite{Werner2013TheSystem}. If the users want to use an existing adapter and need to configure something not supported via the existing adapter, it is hard to get done. Also, there are some cases it is hard to develop an adapter or running alongside the \acrshort{fews}. As an example, the FLO2D model is using for hydrologic modeling and the model only supports only on Windows based operating systems. If you setup the system on Linux based operating system and using Linux based models, it is hard for users to come up with a solution for integrating the FLO2D models.
The \acrshort{fews} support creating workflow process and allow users to run multiple models parallel or run those sequentially as needed, but the scope of the \acrshort{wdias} is not focused on adding such a feature in the system, rather user has to use a scientific workflow management system or come up with own version of workflow.
